The Impact of Immutability on Class Size

### 2020-04-09 paper/article.tex
github action

No similar research has been done so far.

@@

The notion of immutability is not as straightforward as it might seem, and many
different definitions of immutability exist~\cite{potanin2013}.

>>>>

Object immutability has been the subject of multiple researches since
the begining of object-oriented programming. Even though, according to~\cite{potanin2013},
``the notion of immutability is not as straightforward as it might seem, and many
different definitions of immutability exist,'' ...

Even though Java encourages programmers to explicitly make objects immutable
by attaching the \texttt{final} modifier to its attributes, very often, as was
demonstrated by~\cite{unkel2008}, the modifier is not used even when
attributes are not modified after object instantiation or, as demonstrated
by~\cite{nelson2012}, after initialization. Such a delayed initialization of attributes, as
explained by~\cite{fahndrich2007}, leads to the prevalence of \texttt{null}
and the ability to initialize circular data structures, while both of these
practices have negative impact on software quality.

Despite the existence of the \texttt{final} modifier, Reflection API enables
modification of any attributes in Java. Some mechanisms were suggested to
overcome this technical vulnerability, for example
via ``freezer'' objects~\cite{leino2008},
a new type system~\cite{tschantz2005},
or immutability assertion framework~\cite{pechtchanski2005}.
However, without any modifications to Java~\cite{steele2005},
thanks to the existence of Reflection API,
it's not possible to say whether a particular Java object is modified
after instantiation or not---any object can be modified.

Aside from the \texttt{final} modifier, which is an explicit way of declaring
read-only status of an attribute, there are also methods of
code analysis, enabling the detection of implicit immutability. For example,
a static flow-sensitive analysis algorithm was introduced by~\cite{porat2000}
to classify fields and classes as either mutable or immutable.

@@

...
